---
title: "Reproducible Research: Peer Assessment 1"
author: "Ashraf"
date: "October 9, 2018"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r, echo = TRUE}
fileName<-"Dataset.zip"
lnk<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

if(!file.exists(fileName))
{
  download.file(lnk,destfile="Dataset.zip", mode = "wb")
}
if(file.exists(getwd())){
  unzip("Dataset.zip", files = NULL, exdir=".")
}
```

##Code for reading in the dataset and/or processing the data
```{r, echo = TRUE}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
summary(activity)
```

##1.What is mean total number of steps taken per day?
##Histogram of the total number of steps taken each day
```{r, echo = TRUE}
known_Steps <-activity[!is.na(activity$steps),]
dailySteps <- aggregate(steps ~ date, data = known_Steps, sum)
hist(as.numeric(dailySteps$steps), breaks = 30, col = "blue", xlab = "# Steps", main= "Total number of steps taken each day")
#mean
mean(dailySteps$steps)
#Median
median(dailySteps$steps)
```


##2.What is the average daily activity pattern?
##Time series plot of the average number of steps taken
```{r, echo = TRUE}
AvgDailyPatern <- aggregate(steps ~ interval, data = known_Steps, mean)

plot(AvgDailyPatern$interval, AvgDailyPatern$steps, type = "l", col="blue", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average Daily Activity Patern")

MaxAvgSteps<-max(AvgDailyPatern$steps)
obsNum<-match(MaxAvgSteps,AvgDailyPatern$steps)


#Maximum Number of Steps
max(AvgDailyPatern$steps)
#The 5-minute interval countain the maximum number of steps
AvgDailyPatern[obsNum,]$interval

```



## Imputing missing values
```{r, echo = TRUE}
#Total number of missing values in the dataset
sum(is.na(activity$steps))

#update the NA value with the mean steps of the interval.
unknown_Steps <-activity[is.na(activity$steps),]
Means_steps <- aggregate(steps ~ interval, data = known_Steps, mean)
#update NA Steps with mean value for each interval
unknown_Steps$steps <- as.numeric(Means_steps$steps)

#New dataset that is equal to the original dataset but with the missing data filled in.
new_activity <- rbind(known_Steps,unknown_Steps)
new_activity<-  new_activity[order(new_activity[,2],new_activity[,3]),]

#Histogram to plot the new dataset
newdailySteps <- aggregate(steps ~ date, data = new_activity, sum)
hist(as.numeric(newdailySteps$steps), breaks = 30, col = "blue", xlab = "# Steps", main= "Total number of steps taken each day")

#New mean
mean(newdailySteps$steps)
#New Median
median(newdailySteps$steps)
```


## Are there differences in activity patterns between weekdays and weekends?
```{r, echo = TRUE}
#Append Days to the new dataset
days <- weekdays(new_activity$date)
new_activity <- cbind(new_activity,days)

#Weekday & weekends
new_activity$Weekday <- ifelse((new_activity$days %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")), TRUE, FALSE)

#Average number of steps for weekdays
weekdaysActivities <- new_activity[new_activity$Weekday,]
weekdaysStepsIntervals <- aggregate(steps ~ interval+Weekday, data = weekdaysActivities, FUN=mean)

#Average number of steps for Weekens
weekendsActivities <- new_activity[!new_activity$Weekday,]
weekendsStepsIntervals <- aggregate(steps ~ interval+Weekday, data = weekendsActivities, FUN=mean)

allaggregatedData<- rbind(weekendsStepsIntervals, weekdaysStepsIntervals)
allaggregatedData$Daytype <- ifelse(allaggregatedData$Weekday == TRUE, "Weekday", "Weekend")


library(lattice)
#Ploting
xyplot(steps ~  interval |Daytype, data = allaggregatedData, layout = c(1,2), type ="l",xlab="Intervals", ylab="# Steps")
```
